---
title: "trafficr"
author: "douglascm"
date: "1/2/2020"
output: github_document
df_print: paged
---

```{r load, include = FALSE, eval = TRUE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
library(trafficr)
```
# Traffic Assignment Package
trafficr is a package for static traffic assignment R code. Currently, the program can solve the static traffic assignment problem using user equilibrium (UE) for the city network. The solution is achieved using the Frank-Wolfe algorithm.

## How to run traffic assignment
Following steps explain the setup and how to run the procedures and extract flows and maps.

### 1. Install the R package with: 
```{r install, eval=FALSE}
devtools::install_github("douglascm/trafficr")
```
Package devtools is required for the installation, install with `install.packages('devtools')`

### 2. Data Preparation : 
Import into the current environment all three required datasets:

- *demand* tabular data containing demand data for each origin and destination:
```{r demand load, eval=FALSE}
demand <- fread("path to file")
```
    - Required columns:
        - orig: origin id (integer)
        - dest: destination id (integer)
```{r demand}
head(sioux_demand)
```
- *network* sf object data containing road network:
```{r network load, eval=FALSE}
network <- st_read("path to file")
```
    - Required columns:
        - id_link: link id (integer)
        - ffs_fw: forward free flow speed in km/h (numeric)
        - ffs_bw: backward free flow speed in km/h (numeric)
        - cap_fw: forward link capacity in veh/h (numeric)
        - cap_bw: backward link capacity in veh/h (numeric)
    - Optional columns:
        - addv_fw: forward additional cost (numeric)
        - addv_bw: backward additional cost (numeric)
        - wg_fw: forward cost weight (numeric)
        - wg_bw: backward cost weight (numeric)
        - alpha: alpha BPR function parameter (numeric)
        - beta: beta BPR function parameter (numeric)
        - cost: actual network cost, used instead of geometric distance and speed (numeric)
```{r network}
head(sioux_network)
```
- *zones* sf object data containing zones to be added to road network:
```{r zones load, eval=FALSE}
zones <- st_read("path to file")
```
    - Required columns:
        - id_zone: link id (integer)
        - longitude: Longitude in WGS84 (numeric)
        - latitude: Latitude in WGS84 (numeric)
```{r zones}
head(sioux_zones)
```

### 3. Running the code :
Use the following methods to perform operations:

`configure_graph()` Tranforms network data from source network and zones and outputs a list with 2 elemens, 1 being the graph and 2 being the transformed zones:
```{r configure graph, eval=FALSE}
configure_graph(network,zones,turn_penalty,left_side,addvalue_cols,weighted,performance_measures,use_cost_col)
```
- *network*: sf object containing network data
- *zones*: sf object containing zone data
- *turn_penalty*: Adds turn penalty (see dodgr documentation on turn penalty from function 'R/weight-streetnet.R')
- *left_side*: Does traffic travel on the left side of the road (TRUE) or the right side (FALSE)? - only has effect on turn angle calculations for edge times. (see dodgr documentation on turn penalty from function 'R/weight-streetnet.R')
- *addvalue_cols*: Does network shapefile has columns 'addv_fw' and 'addv_bw' for added cost in links? (TRUE) or (FALSE)
- *weighted*: Does network shapefile has columns 'wg_fw' and 'wg_bw' for a cost multiplication factor in links? (TRUE) or (FALSE)
- *performance_measures*: Does network shapefile has columns 'alpha' and 'beta' for usage on BPR funcion? (TRUE) or (FALSE)
- *use_cost_col*: Use cost column as cost (opposite to speed and distance)? (TRUE) or (FALSE)

`configure_demand()` Tranforms demand data from source origin and destination to match origin and destination from graph/zone objects:
```{r configure demand, eval=FALSE}
configure_demand(demand,zones)
```
- *demand*: demand data
- *zones*: zones object created with configure_graph()

`solve_ue()` Solve the traffic flow assignment model (user equilibrium) by Frank-Wolfe algorithm (all the necessary data must be  properly input into the model in advance).
```{r solve ue, eval=FALSE}
solve_ue(graph,demand,col,max_iterations)
```
- *graph*: Graph object created with configure_graph() function
- *demand*: Demand object created with configure_demand() function
- *col*: Column from demand file selected for assignment
- *max_iterations*: Maximum number of iterations for Frank-Wolfe algorithm

Next is an example of the structure to run the entire code, with additional functions that are documented inside R:
```{r example}
flist <- configure_graph(sioux_network,sioux_zones,use_cost_col=TRUE)

graph <- flist[[1]]
zones <- flist[[2]]
demand <- configure_demand(sioux_demand,zones)

graph <- solve_ue(graph,demand,col = 'demand')
times <- get_times(graph,demand)
paths <- get_paths(graph,demand)

network <- graph_to_sf(graph,sioux_network)
plot_flowmap(network,show_labels = T)
```

## Questions
Feel free to send an email to [douglas.capelossi@gmail.com](douglas.capelossi@gmail.com) if you have questions or concerns.


## Future releases
Future releases will have an implementation of other traffic assignment algorithms such as Gradient Projection, Origin-based assignment, and Algorithm B.
